### 2025-08-29 sections/1-introduction.tex
back down to 10 pages with some extra material

Several algorithms have been developed to tackle $k$-nearest neighbor search on large datasets, including Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
However, many of these algorithms do not support \emph{exact} search (as defined in Section \ref{sec:introduction}).

Hierarchical Navigable Small World networks~\cite{malkov2016hnsw} rely on navigable small world (NSW) networks and skip lists.
InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
ANNOY~\cite{annoy} uses random projection and tree building.
At each intermediate node of the tree, a random hyperplane, defined by two sampled points, splits the space.
Increasing the number of such trees improves recall but increases search time.

@@

This paper introduces CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in Rust.
We provide a theoretical analysis of their time complexity in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a uniform distribution.
Given that CAKES's algorithms are intended for datasets with a manifold structure, complexity analyses assuming a uniform distribution would be uninformative.
As a result, our analysis assumes datasets with manifold structure—characterized by low fractal dimension and metric entropy—reflecting the conditions for which CAKES is designed.

>>>>

Several algorithms have been developed to tackle $k$-nearest neighbor search on large datasets, including Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.

HNSW~\cite{malkov2016hnsw} relies on navigable small world (NSW) networks and skip lists.
IVF~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
ANNOY~\cite{annoy} uses random projection and tree building.
However, many of these algorithms do not support \emph{exact} search (as defined in Section \ref{sec:introduction}).

@@

This paper introduces CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in Rust.
We provide a theoretical complexity analysis in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a uniform distribution.
Given that CAKES's algorithms are intended for datasets with a manifold structure, an analysis assuming a uniform distribution would be uninformative.
As a result, our analysis assumes datasets with manifold structure, reflecting the conditions for which CAKES is designed.
